### 2014-01-28 limits/ms.tex
Justin Changes

\documentclass[galley,usenatbib]{mn2e}
%\documentclass[twocolumn,galley]{mn2e}
%\documentclass[onecolumn,galley,draft]{mn2e}
\usepackage{myaasmacros}
\usepackage{mathrsfs}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{ulem} % Only necessary for striking out TODO items see \sout
\usepackage{color} % Only necessary for hilight

\newcommand{\Glass}{{\sc Glass}}
\newcommand{\PixeLens}{{\sc PixeLens}}
\newcommand{\Rmap}{\ensuremath{R_\mathrm{map}}}
\newcommand{\Rpix}{\ensuremath{R_\mathrm{pix}}}
\newcommand{\M}{\ensuremath{\mathscr{M}}}
\newcommand{\E}{\ensuremath{\mathscr{E}}}
\newcommand{\eps}{\ensuremath{\varepsilon}}

\newcommand{\Eavg}{\ensuremath{\langle \E \rangle}}

\newcommand{\hilight}[1]{{\bf \color{red} #1}}

\newcommand{\kpc}{\ensuremath{\mathrm{kpc}}}
\newcommand{\Msun}{\ensuremath{\mathrm{M}_\odot}}
\newcommand{\tabref}[1] {Table~\ref{#1}}
\newcommand{\figref}[1] {Figure~\ref{#1}}
\newcommand{\eqnref}[1] {Eq.~(\ref{#1})}
\newcommand{\eqnrefp}[1] {(Eq.~\ref{#1})}
\newcommand{\secref}[1] {\S\ref{#1}}
\newcommand{\appref}[1] {Appendix~\ref{#1}}
\newcommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\renewcommand{\vec}[1]{\ensuremath{\boldsymbol{#1}}}

\newcommand{\mockAA}{{\sc star1.0-dmCore}}
\newcommand{\mockAC}{{\sc star1.0-dmCusp}}
\newcommand{\mockBB}{{\sc star1.5-dmCore}}
\newcommand{\mockBC}{{\sc star1.5-dmCusp}}

%For deprojection algorithm: 
\def\cth{\cos\theta}
\def\sth{\sin\theta}
\def\dth{{\rm d}\theta}

% From aastex.cls
\newcommand\plotone[1]{%
 \centering
 \leavevmode
 \includegraphics[width={\columnwidth}]{#1}%
}%
\newcommand\plottwo[2]{{%
 \centering
 \leavevmode
 \columnwidth=.45\columnwidth
 \includegraphics[width={\columnwidth}]{#1}%
 \hfil
 \includegraphics[width={\columnwidth}]{#2}%
}}%
\newcommand\plotthree[3]{{%
 \centering
 \leavevmode
 \columnwidth=.30\textwidth
 \includegraphics[width={\columnwidth}]{#1}%
 \hfil
 \includegraphics[width={\columnwidth}]{#2}%
 \hfil
 \includegraphics[width={\columnwidth}]{#3}%
}}%

\title[Lens Recovery with \Glass]{Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens}
\author[J.~P.~Coles~et~al.]{%
Jonathan P.~Coles,$^{1,3}$\thanks{jonathan@exascale-computing.eu}%
\newauthor%
Justin I.~Read$^2$,%
\newauthor%
and Prasenjit Saha$^3$%
%
\\
$^1$Exascale Research Computing Lab, Universit\'e de Versailles Saint-Quentin-en-Yvelines,\\$\quad$55 Avenue des Etats Unis, 78000, Versailles, France\\
$^2$\hilight{Department of Physics, University of Surrey, Guildford, Surrey, GU2 7XH, United Kingdom} \\
$^3$\hilight{Institute of Physics, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland}
}

\begin{document}
\maketitle

\begin{abstract}
We use a new non-parametric gravitational modelling tool --- \Glass\ --- to
determine what quality of data (strong lensing, stellar kinematics, and/or
stellar masses) are required to measure the circularly averaged mass profile of
a lens and its shape. \Glass\ uses an under-constrained adaptive grid of mass
pixels to model the lens, searching through thousands of models to marginalise
over model uncertainties. Our key findings are as follows: (i) for pure lens
data, multiple sources with wide redshift separation give the strongest
constraints as this breaks the well-known mass-sheet or steepness degeneracy;
(ii) a single quad with time delays also performs well, giving a good recovery
of both the mass profile and its shape; (iii) stellar masses -- for lenses
where the stars dominate the central potential -- can also break the steepness
degeneracy, giving a recovery almost as good as having time delay data or
multiple source redshifts; (iv) stellar kinematics provide a robust measure of the mass at the half light radius of the stars $r_{1/2}$ that can also break the steepness degeneracy if $r_{1/2} \neq r_E$ -- the Einstein radius; and (v) if $r_E \sim r_{1/2}$, then stellar kinematic data can be used to probe the stellar velocity anisotropy 
$\beta$ -- an interesting quantity in its own right. Where
information on the mass distribution from lensing and/or other probes becomes
redundant, this opens up the possibility of using strong lensing to constrain
cosmological models. We will study this, and present the first results from
\Glass\ applied to real data, in forthcoming papers.
\end{abstract}

\section{Introduction}\label{sec:intro}

Strong gravitational lenses are rare. Since the discovery of the first lens Q0957+561 \citep{1979Natur.279..381W}, just $\sim400$ have been discovered to date\footnote{See e.g. {\tt http://www.masterlens.org} for a catalogue.}. However, this number is expected to increase to
several thousand over the next ten years as new surveys, both
ground-based\footnote{http://pan-starrs.ifa.hawaii.edu}$^{\rm,}$\footnote{http://www.darkenergysurvey.org}
and space-based\footnote{http://www.euclid-ec.org} -- together with a
community of citizen-science volunteers examining the image data for
candidates\footnote{{\tt http://spacewarps.org}} -- come online.

Since lensing depends only on gravity, strong lenses offer a
unique window onto dark matter and cosmology \citep{2010CQGra..27w3001B,2012arXiv1206.1225A}. However, 
extracting dark matter properties or cosmological constraints from these lensing data will require sophisticated modelling. In particular, with an unprecedented data set imminent, it is prudent to look again at systematic errors in the lens models to determine what quality of data (in particular complementary data from stellar/gas kinematics, lens time delays and/or stellar mass constraints) are required to address problems of interest. It is towards that goal that this present work is directed.

To see why lens modelling details are of crucial importance, let us recall the
essential quantities that appear in lensing.  First we have the
distances. Let $D_L$, $D_S$, $D_{LS}$ be the angular-diameter
distances to the lens, source, and from lens to source; these are all proportional to $c/H_0$ (where $c$ is the speed of light in vacuo, and $H_0$ the Hubble constant), with additional terms that depend on the particular choice of cosmology \citep[e.g.][]{1999coph.book.....P}. Typically:
\begin{equation}
D_L \approx z \frac c{H_0}\ \hbox{and}\ \frac{D_S}{D_{LS}} \sim 1.
\end{equation}

For multiple images, the sky-projected density must exceed the critical
density \citep[e.g.][]{1992grle.book.....S}:

\begin{equation}
\Sigma_{\rm crit} = \frac{D_S}{D_{LS}} \Sigma_L = \frac{D_S}{D_{LS}} \frac{c^2}{4\pi G D_L} \sim \frac{1\rm\,kg\,m^{-2}}{z_L}
\end{equation}
in some region.  The separation between the lensed images is of order
the Einstein radius $\theta_E$, which is related to
the mass by:
\begin{equation}
\theta_E \sim \frac{R_G}{D_L} \frac{D_{LS}}{D_S}
\end{equation}
where $R_G = GM/c^2$ (with $M$ the projected mass enclosed within $\theta_E$) is the gravitational radius. If the source is a
quasar or otherwise rapidly variable, a time delay $\Delta t$ in the variability
will be present where:
\begin{equation}
\Delta t \sim R_G/c \,.
\end{equation}
So in principle, one can not only measure the mass of the lens, one
can use the dependence on the cosmology-dependent $D$ factors to
extract the cosmological model and all its parameters.
\cite{1937ApJ....86..217Z} drew attention to the former, and
\cite{1964MNRAS.128..307R,1966MNRAS.132..101R} pointed out the latter,
all long before lenses were discovered.  The difficulty with actually doing this, however, became
apparent soon after the discovery of the first lens by
\cite{1979Natur.279..381W}.  In the first ever paper on lens
modelling, \cite{1981ApJ...244..736Y} found that many plausible mass
distributions could reproduce the data.

\cite{1981ApJ...244..736Y} were remarkably prescient about the
subsequent development of lens modelling.  First, they introduced the
technique of choosing a parametric form for the lensing mass and then
fitting for the parameters, which is still the most common strategy
\citep[see for example][]{2010GReGr..42.2151K,2011A&ARv..19...47K}.
Second, they pointed out the non-uniqueness of lens models -- lensing
degeneracies.  Third, they suggested combining lensing data with
stellar kinematics and X-rays, to reduce the effect of the
degeneracies.  Later work, as well as following up these suggestions,
has introduced some further new ideas.  Four of these are important for the present
work:
\begin{enumerate}
\item {\bf Free-form modelling:} In `free-form' or non-parametric modelling, there is no specified parametric form for
  the mass distribution.  There are still assumptions (or priors) on
  the mass distribution, such as smoothness or being centrally
  concentrated
  \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2009A&A...500..681M,2010ApJ...723.1678C}
  but these are much less restrictive than parametric forms (\cite{2006MNRAS.367.1209L} implemented a particularly elegant prior, which is that the mass distribution must be non-negative and no
  extra images are produced). To be concrete, we define from here on:

  \begin{quote}
  {\it Non-parametric, or `free-form' $\equiv$ more parameters than data constraints (i.e. deliberately under-constrained)}
  \end{quote}

  Being under-constrained, it is then {\it necessary} to explore model degeneracies rather than find a single `best-fit' solution. Free-form models are more commonly used with
  cluster lenses \citep{2006ApJ...652L...5S,2009ApJ...690..154S,2009A&A...500..681M,2013arXiv1304.2393S}, but
  can be used with galaxy lenses as well, where their less restrictive
  assumptions can be important.  For example, in time-delay galaxy
  lenses, parametric models have historically led to tension with independent measures of the Hubble parameter
  $H_0$ \citep[e.g.][]{2002astro.ph..4043K,2002ApJ...578...25K}; these are resolved once the less restrictive assumptions of free-form models are permitted \citep{2007ApJ...667..645R}.
\item Model ensembles, exploring a diverse range of possible mass
  distributions that nonetheless all fit the data, are a way of
  combating the non-uniqueness of models.  Such ensembles are
  possible in parametric models \citep{1999AJ....118...14B} but are
  more common in free-form models
  \citep{2000AJ....119..439W,2009ApJ...690..154S,2012MNRAS.425.3077L}.
\item {\bf Stellar kinematic constraints:} This was first suggested by \citet{2002MNRAS.337L...6T} as a means to break lensing degeneracies and determine cosmological parameters. This works because stellar kinematics can provide an independent estimate of the Einstein radius, via the virial theorem:
%
\begin{equation}
  \frac{\langle v^2_{\rm los}\rangle}{c^2} \approx
  \frac{\theta_E}{6\pi} \frac{D_S}{D_{LS}}
\label{eqn:virial}
\end{equation}
 where $\langle v^2_{\rm los}\rangle$ is the projected stellar velocity dispersion, and the above relation becomes exact for isothermal lenses.  The use of
  two-dimensional kinematics by \cite{2011MNRAS.415.2215B} is
  especially interesting.
\item {\bf Stellar mass constraints:} The stellar mass in a lens can be inferred from photometry and
  compared with the total mass
  \citep{2005ApJ...623L...5F,2008MNRAS.383..857F,2011ApJ...740...97L}.
  Since the inferred stellar mass depends on the assumed IMF,
  lenses in which stellar mass dominates can be used to derive upper
  bounds on the stellar $M/L$ \citep{2010MNRAS.409L..30F}. Lower bounds
  on stellar $M/L$ have also been claimed \citep{2013MNRAS.428.3183D}
  but these require upper-bound assumptions about the dark-matter
  profile. \hilight{JR. I feel like we need more citations here and throughout. We must be skipping over an Armada of literature.}
\item {\bf Testing modelling strategies:} Using mock data to see how well a given model can recover
  simulated lenses is increasingly being recognised as essential.  Simple
  blind tests have appeared in earlier work \citep[for example, Figure
    2 in][]{2000AJ....119..439W}, but more recently, tests against
  dynamically simulated galaxies or clusters are favoured
  \citep{2007ApJ...667..645R,2007MNRAS.380.1729L,2009A&A...500..681M,2009MNRAS.393.1114B,2010ApJ...723.1678C}.
\end{enumerate}

There are two further key modelling ideas in the literature that we will not touch upon in this present work: to use
X-ray intensity and temperature profiles as a mass constraint
\citep[e.g.][]{2013ApJ...765...25N}; and to model
multiple lenses simultaneously, with one or more cosmological
parameters variable but shared between the lenses. This latter strategy has
been used to constrain $H_0$ from time delay lenses 
\citep{2006ApJ...652L...5S,2008ApJ...679...17C,2010ApJ...712.1378P}.
(In principle, cosmological parameters can be constrained even from a
single lens if the data are good enough. Some recent models have attempted this starting from a prior
based on the Cosmic Microwave Background (CMB) radiation \citep{2010Sci...329..924J,2013arXiv1306.4732S}.)

In this paper, we use a new program --- \Glass\ (Gravitational Lensing AnalysiS Software) --- to determine what
combination of lensing, stellar mass, and stellar kinematic
constraints best constrain the projected mass profile and shape of several
gravitational lens produced from mock data.  The compute-intensive core of \Glass\ is a
parallel implementation of the high-dimensional sampling algorithm
from \cite{2012MNRAS.425.3077L}.  Around that core, \Glass\ has
extensible code (in software terminology, a framework) in Python,
encoding the lens equation, a basis and priors for the mass
distribution, and various post-processing and visualisation modules.  \Glass\
is described in \secref{sec:glass}.  In \secref{sec:theory}, we
review the key elements of lensing theory, stellar population synthesis, and
stellar dynamics we will need. In \secref{sec:mockdata}, we describe our mock data. In \secref{sec:results}, we present our results from applying \Glass\ to these mock data. Finally, in \S\ref{sec:conclusions} we present our conclusions.

\section{Theory}\label{sec:theory}

\subsection{Lensing essentials}\label{sec:lensing_basic}

We follow \cite{1986ApJ...310..568B} with minor differences in notation.

The lens equation:
\begin{equation}
\begin{aligned}
\vec\beta &= \vec\theta - \frac{D_{LS}}{D_S}\vec\alpha(\vec\theta) \\
\vec\alpha(\vec\theta) &= \frac{4G}{c^2D_L} \int \Sigma(\vec\theta')
                          \frac{(\vec\theta - \vec\theta')}
                          {\ |\vec\theta - \vec\theta'|^2} \, d^2\vec\theta'
\end{aligned}
\label{eqn:lens_equation}
\end{equation}
maps an observed image position $\vec\theta$ to a source position
$\vec\beta$.  The lens can be thought of as a projected surface
density $\Sigma$ which diverts the path of a photon instantaneously
through the bending angle $\vec\alpha$.  The $D$ factors, as in the previous section, are angular
diameter distances, which depend on the particular choice of cosmological
parameters $\Omega$:
\begin{equation}
D_{LS} = \frac c{H_0} \frac{1+z_S}{1+z_L} \int_{z_L}^{z_S}
                      \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
%
and $D_L \equiv D_{0,L}$, $D_S \equiv D_{0,S}$.
One way to understand the lens equation is via Fermat's principle. We
can think of light as travelling only along extremum paths where
lensed images occur. Such paths occur at
the extrema of the photon {\it arrival time\/} $t(\vec\theta)$
that depends on the geometric path the photon takes and the general
relativistic gravitational time dilation due to a thin lens at
redshift $z_L$:
\begin{equation}
\begin{aligned}
\frac{ct(\vec\theta)}{(1+z_L)D_{L}}
&= {\textstyle\frac12} |\vec\theta - \vec\beta|^2
   \times \frac{D_{S}}{D_{LS}} \\
&- \frac{4GD_L}{c^2}
   \int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| \, d^2\vec\theta'
\label{full arrival time}
\end{aligned}
\end{equation}
We can simplify the above equation by introducing a dimensionless time $\tau$ and density $\kappa$:

\begin{equation}
\tau(\vec\theta) = \frac{ct(\vec\theta)}{(1+z_L)D_{L}} \,\,\,;\,\,\, \kappa(\vec\theta) \equiv \frac{\Sigma(\vec\theta)}{\Sigma_L}
\end{equation}
and hence rewrite \eqnref{full arrival time} as:
\begin{equation}
\tau(\vec\theta) = {\textstyle\frac12} |\vec\theta - \vec\beta|^2
                   \times \frac{D_{S}}{D_{LS}}
                 - \frac1\pi \int \kappa (\vec\theta')
                   \ln|\vec\theta - \vec\theta'| d^2\vec\theta'
\label{arrival time}
\end{equation}
The scaled arrival time $\tau$ is like a solid angle. It is of order
the area (in steradians) of the full lensing system. The expression
$|\vec\theta - \vec\beta|^2$ is of order the image-separation
squared, and the other terms are of similar size.  For this reason, is
convenient to measure $\tau$ in arcsec$^{2}$.

Lensing observations provide information only at $\vec\theta$ where
there are images.  Hence, the arrival-time surface $\tau(\vec\theta)$
is not itself observable.  Its usefulness lies in that observables can
be derived from it.  An image observed at $\vec\theta_1$ implies
that $\nabla\tau(\vec\theta_1)=0$.  A measurement of time delays
between images at $\theta_1$ and $\theta_2$ implies that
$t(\vec\theta_1)-t(\vec\theta_2)$ is known.  Interestingly, both these
types of observations give contraints that are linear in $\kappa$ and
$\vec\beta$.

The rather complicated dependence of lensing observables on the mass
distribution $\kappa(\vec\theta)$ has an important consequence: very different
mass distributions can result in similar observables.  This is the phenomenon
of lensing degeneracies.  While the non-uniqueness of lens models noted by
\cite{1981ApJ...244..736Y} already hinted at degeneracies, their existence was
first derived by \cite{1985ApJ...289L...1F}.  The most important is the
so-called mass-sheet degeneracy, which is that image positions remain invariant
if $\tau(\vec\theta)$ is multiplied by an
arbitrary constant.  This corresponds to rescaling the surface density at the images $\kappa(\vec\theta)$. In fact there are infinitely many degeneracies
\citep{2000AJ....120.1654S} because any transformation of the arrival-time
surface away from the images has no effect on the lensing observables.  In
particular, there are degeneracies that involve the shape of the mass
distribution \citep{2006ApJ...653..936S,2013arXiv1306.4675S}.  Degeneracies
tend to be suppressed if there are sources at very different redshifts or
`redshift contrast' \citep{1998AJ....116.1541A,2009ApJ...690..154S}, because
the presence of different factors of $D_S/D_{LS}$ in the image plane makes it
more difficult to change the mass distribution and the arrival-time surface
without affecting the lensing observables.  But degeneracies are still present
with multiple source redshifts.  \citep{2008MNRAS.386..307L}.

\subsection{Stellar populations}

For many galaxy lenses, the gravitational potential in the inner
region is dominated by the stellar mass.  Steller mass can be
estimated by combining photometry and colours with models of the
stellar populations.  Such estimates are reasonably robust, even if
the star-formation history is very uncertain: given a
stellar-population model \citep[such as][]{2003MNRAS.344.1000B} and an
initial mass function (IMF), the stellar mass can be inferred to 0.1
to 0.2 dex using just two photometric bands \citep[see e.g., Figure~1
  in][]{2008MNRAS.383..857F}.  By comparing the lensing-mass and
stellar-mass profiles in elliptical galaxies, it is possible to
extract the radial dependence of the baryonic vs dark-matter fraction
\citep{2005ApJ...623L...5F,2008MNRAS.383..857F,2011ApJ...740...97L}.

The major uncertainty at present in the stellar mass is probably the
IMF.  In the lensing galaxy of the Einstein Cross, the IMF cannot be
much more bottom-heavy than \cite{2003PASP..115..763C}, because
otherwise the stellar mass would exceed the lensing mass
\cite{2010MNRAS.409L..30F}.  More massive galaxies, however, do appear
to have more of their stellar mass in low-mass stars.  This is
indicated by molecular spectral features characteristic of low mass
stars
\citep{2004ApJ...614L.101C,2012ApJ...747...69C,2013MNRAS.429L..15F}.
The \cite{2003PASP..115..763C} IMF would, however, still provide a robust
lower limit on the stellar mass and hence, also a limit on the total
mass.  Accordingly, \Glass\ allows a constraint of the form
\begin{equation} 
M(\vec\theta) \geq M_{\rm stel}(\vec\theta)
\end{equation} 
on the total mass.

\subsection{Stellar kinematics}\label{sec:kinematics}

Another useful constraint follows from the velocity of stars within the lensing
galaxy. Assuming spherical symmetry, stars obey the projected Jeans equations
\citep[e.g.][]{2008gady.book.....B}:

\begin{equation}
\sigma_p^2(R) = \frac{2}{I(R)}\int_R^\infty dr \left(1-\beta \frac{R^2}{r^2}\right) \frac{\nu \sigma_r^2 r}{\sqrt{r^2 - R^2}};
\label{eqn:sphericaljeans}
\end{equation}
\begin{equation} 
\sigma_r^2(r) = \frac{r^{-2\beta}}{\nu}\int_r^\infty r'^{2\beta} \nu \frac{GM_{\rm 3D}(r')}{r'^2}dr'
\end{equation} 
where $\sigma_p$ is the projected velocity dispersion of the stars as a
function of projected radius $R$; $I(R)$ is the surface density of the stars;
$\nu(r)$ is the three dimensional stellar density; $\sigma_{r,t}(r)$ are the
radial and tangential velocity dispersions, respectively; $\beta(r) = 1 -
\sigma_t^2/2\sigma_r^2 = \mathrm{const.}$ is the velocity anisotropy (here
assumed to be constant, and not to be confused with $\vec\beta(\vec\theta)$ from
lensing); $G$ is Newton's gravitational constant; and $M_{\rm 3D}(r)$ is
the mass profile that we would like to measure. By convension, we always write $R$ 
for a projected radius, and $r$ for a 3D radius.

It is immediately clear from \eqnref{eqn:sphericaljeans} that, even assuming
spherical symmetry, we have a degeneracy between the enclosed mass profile
$M_{\rm 3D}(r)$ and the velocity anisotropy $\beta(r)$. This can be understood
intuitively since $\beta(r)$ measures the relative importance of radial versus
circular orbits and is intrinsically difficult to constrain given only one
component of the velocity vector for each star. Nonetheless, $\beta(r)$ can be
constrained given sufficiently many stars, since radial Doppler velocities
sample eccentric orbits as $r\rightarrow 0$ and tangential orbits as
$r\rightarrow \infty$ \citep[e.g.][]{2002MNRAS.330..778W}. It can also be
estimated if an independent measure of $M_{\rm 3D}(r)$ is available -- for example
coming from strong lensing.

While $M_{\rm 3D}(r)$ is difficult to measure from stellar kinematics alone, the mass within the half light radius is robustly recovered \citep[e.g.][]{2009ApJ...704.1274W,2010MNRAS.406.1220W,2012ApJ...754L..39A} since stellar systems in dynamic quasi-equilibrium obey the virial theorem (equation \ref{eqn:virial}). This means that stellar kinematics can break the steepness degeneracy if $r_{1/2} \neq r_E$, where $r_E = D_L \theta_E$ is the physical Einstein radius. We test this expectation in \S\ref{sec:results}.

We describe our numerical solution of \eqnref{eqn:sphericaljeans} in
\S\ref{sec:glasskinematics} and present tests applied to mock data in
\S\ref{sec:results}.

\section{Numerical Methods}\label{sec:glass}

\subsection{A new lens modelling framework: \Glass}

\Glass\ is the Gravitational Lensing AnalysiS Software. It is an evolution of the concepts employed by the free form modelling tool
\PixeLens\ but completely rewritten in Python and C. 
The most compute intensive
portion was written in C but Python was chosen because of its flexibility as a 
language and for its large scientific library support. The flexibility allows 
\Glass\ to have quite sophisticated behavior while at the same time simplifying the 
user experience and reducing the overall development time. One of the striking
features is that the input file to \Glass\ is itself a Python program.
Understanding Python is not necessary for the most basic use, but this allows
a user to build complex analysis of a model directly into the input file. \Glass\
may also be used as an external library to other Python programs.  The software
is freely available for download or from the author%
%
\footnote{http://www.jpcoles.com.}.

The key scientific and technical improvements are:
%
\begin{enumerate}
  \setcounter{enumi}{0}
  \item A modular framework that allows new priors to be added and modified easily.
\end{enumerate}
%
Each prior is a simple function that adds linear constraints that operate on either
a single lens object or the entire ensemble of objects. \Glass\ comes with a number
of useful priors, the default ones will be described later, but a user can
write their own directly in the input file, or by modifying the source code.
%
\begin{enumerate}
  \setcounter{enumi}{1}
  \item The basis functions approximating a model can be changed. 
\end{enumerate}
%
\Glass\ currently describes the lens mass as a collection of pixels, but the code
has been designed to support alternative methods. In particular, there are future
plans to develop a module using Bessel functions. This will require a new set of 
priors that operate on these functions.
%
\begin{enumerate}
  \setcounter{enumi}{2}
  \item Non-linear constraints can be imposed in an automated post-processing step. 
\end{enumerate}
%
Once \Glass\ has generated an ensemble of models given the linear constraints, any number
of post processing functions can be applied. Not only can these functions be used to
derive new quantities from the mass models, they can also be used as a filter to 
accept or reject a model based on some non-linear constraint. The plotting functions
within \Glass\ will correctly display models that have been accepted or rejected.
%
\begin{enumerate}
  \setcounter{enumi}{3}
  \item The central region can have a higher resolution to capture steep models. 
\end{enumerate}
%
With the default basis set of pixels, the mass distribution of the lens is
described by a uniform grid. However, in the central region of a lensing galaxy
where the mass profile may rise steeply, the center pixel uses a higher
resolution. This allows the density to increase smoothly but still allow for
a large degree of freedom within the inner region without allowing the density
to be arbitrarily high. 
%
\begin{enumerate}
  \setcounter{enumi}{4}
  \item Stellar density can be used as an additional constraint.  
\end{enumerate}
%
The mass in inner regions of galaxies is often dominated by the stellar component
which one can estimate using standard mass-to-light models. This data can be added
to the potential as described later in \secref{stellar mass}. By using the stellar
mass one can place a lower bound on the mass and help constrain the inner most
mass profile.
%
\begin{enumerate}
  \setcounter{enumi}{5}
  \item Point or extended mass objects can be placed in the field.
\end{enumerate}
%
\PixeLens\ only allows for a shear term to be added to the potential (as shown
later in \eqnref{shear}) to account for mass external to the modelled region. This
is useful to capture the gross effects of a distant neighbour. \Glass\ also
has a shear term but additionally allows for analytic potentials to
be included. This can be used to model substructure or multiple neighbors close
to the main lens. The substructure may have only a small effect if the lens is
a single galaxy, but if the lens is a group then a potential can be added for
each of the galaxies. A few standard functions are already included in \Glass\
including those for a point mass, a power law distribution, or an isothermal (a
particular case of the power law).
%
\begin{enumerate}
  \setcounter{enumi}{5}
  \item A new uniform sampling algorithm for high dimensional spaces.
\end{enumerate}
%
At the heart of \Glass\ lies a new algorithm for sampling the high dimensional
linear space that represents the modelling solution space. This algorithm was
described and tested in \cite{2012MNRAS.425.3077L}; it is 
multi-threaded allowing it to run efficiently on many-cored machines.

\subsection{Analysis Tools}\label{sec:tools}
\Glass\ is not only a modeling tool but also an analysis engine. \Glass\
provides many functions for viewing and manipulating the computed models.
These functions can either be called from a program written by the user or by
using the program \textsc{viewstate.py} included with \Glass. There is also a
tool, \textsc{lenspick.py} for creating a lens, either analytically or from an
$N$-body simulation file. To load the simulation data, \Glass\ relies on the
\textsc{Pynbody} library \citep{pynbody} and can thus load any file supported by that package.

\subsection{Discrete models}\label{sec:discrete}
For this paper, we will restrict ourselves to using the pixelated basis set that
has been employed in previous work and is implemented in \PixeLens. Here we
will briefly review the main ideas. The algorithm for generating models in
\Glass\ samples a convex polytope in a high dimensional space whose interior
points are solutions to the lens equation and satisfy other physically
motivated linear priors (the sampling algorithm in \Glass\ is \emph{not} the same as
employed in \PixeLens).  We therefore formulate all of our equations as
equations linear in the unknowns. We describe the density distribution $\kappa$
as a set of discrete grid cells or pixels $\kappa_i$ and rewrite the potential
\eqnrefp{lensing potential} as
%
\begin{equation}
  \psi(\vec\theta) = \sum_n \kappa_n Q_n(\vec\theta)
  \label{discrete potential}
\end{equation}
%
where the sum runs over all the pixels and $Q_n$ is the integral of the logarithm
over pixel $n$. The exact form for $Q$ is described in \appref{Q derivation}.
We can find the discretized lens equation by simply taking the gradient of the
above equations.

The pixels only cover a finite circular area with physical radius $\Rmap$ and
pixel radius $\Rpix$ with the central cell centered on the lensing galaxy. To account for any global
shearing outside this region from, e.g., a neighboring galaxy, we also add to
\eqnref{discrete potential} two shearing terms
%
\begin{equation}
\label{shear}
\gamma_1(\theta_x^2 - \theta_y^2) + 2\gamma_2\theta_x\theta_y\quad.
\end{equation}
%
We can continue adding terms to account for other potentials. For instance,
we may want to impose a base potential over the field, or add potentials
from the presence of other galaxies in the field. \Glass\
already includes potentials for a point mass or an exponential form, but custom
potentials are straightforward to add and can be included directly in the input file.
If the stellar density $\kappa_s$ has been estimated we can use this as a lower bound
where the stellar potential is a known constant of the form \eqnref{discrete
potential}, e.g., $\kappa_n = \kappa_{\mathrm{dm},n} + \kappa_{s,n}$ for a two-component model.

Unfortunately, the lens equation and the arrival times alone are not enough to form a
closed volume in the solution space. We have not imposed any strong constraints
on the density even so much as to ensure that it is non-negative everywhere. We
must therefore provide additional linear constraints so that we only sample
physical models. The explicit implementation for these priors has been
explained in \cite{2008ApJ...679...17C}, but can be summarized as
\begin{enumerate}
\item The density must be non-negative everywhere.
\item The density profile must have a slope everywhere $\le 0$.
\item The local gradient everywhere must point within $45^{\circ}$ of the center.
\item Density variations must be smooth.
\item Image parity is enforced.
\item The density is radially symmetric. (Optional)
\end{enumerate}

Prior (ii) is necessary to prevent rings from forming where the inner region is
less dense. This would be unusual for a galaxy or a cluster. Prior (vi)
enforces a radial symmetry that requires radially opposed pixels (i.e., on
opposite sides of the center) to be equal. This prior is particularly important
for modelling doubles which must be symmetric. If a lens is suspected to be
symmetric this can also significantly improve the modelling time for quads.

In the simplest form, a single model for a lens is a tuple $\M = (\vec\kappa,
\vec\beta, \gamma_1, \gamma_2)$. A single model represents a single point in the
solution space polytope. Using the MCMC sampling strategy described in \cite{2012MNRAS.425.3077L}
we uniformly sample this space. Collectively, the sampled models are referred
to as an ensemble $\E = \{\M_i\}$, where we usually generate $|\E| \sim 1000$ models.

\subsection{Raytracing}\label{Raytracing}
\Glass\ can also determine the position of images and time delays from 
particle-based simulation output given a source position $\vec\beta$. This is
used to generate the lens configurations used in the parameter study.  The
particles are first projected onto a very high resolution grid representing the
lens plane. The centers $\vec\theta_i$ of each of the grid cells are mapped
back onto the source plane using \eqnref{eqn:lens_equation}. If the location on
the source plane $\vec\beta_i$ is within a user specified
$\eps_\mathrm{accept}$ of $\vec\beta$ then $\vec\theta_i$ is 
accepted and further refined using a root finding algorithm until the distance
to $\vec\beta$ is nearly zero. If multiple points converge to an
$\eps_\mathrm{root}$ of each other then only one point is taken.  Care must be
taken that the grid resolution is high enough that the resulting image position
error is below the equivalent observational error. Time delays are then
calculated in order of the arrival time at each image \eqnrefp{tau}.

\subsection{Removing models with extra images}\label{sec:glassextraimages} 
While linear constraints are applied in \Glass\ by the nature of the sampling
algorithm, non-linear constraints must be applied in post-processing. Models
that are inconsistent with such constraints must then be statistically
discarded via a likelihood analysis. An example of such a non-linear constraint
is the spurious presence of unobserved images. This `null-space' prior was
first proposed and explored by \citet{2006MNRAS.367.1209L} and found to be
extremely powerful. We find that our gradient prior in \Glass\ (see
\S\ref{sec:discrete}), performs much of the same function as
\citeauthor{2006MNRAS.367.1209L}'s null-space prior, but some models can still
rarely turn up spurious images. We reject these in a post-processing step,
where we sweep through the model ensemble applying the ray tracing algorithm
described in \ref{Raytracing}. \hilight{JR. Typically, this results in a rejection rate of XXX\% IS this working? Can we run it once and add the number here?}.

\subsection{A post-processing module for stellar kinematics}\label{sec:glasskinematics} 
Similarly to the null-space constraint (\S\ref{sec:glassextraimages}), stellar
kinematic constraints constitute a non-linear prior on the mass map and must be
applied in post-processing. We sweep through the model ensemble performing an
Abel deprojection to determine $M_{\rm 3D}(r)$ from the projected surface density $\Sigma(R)$
assuming spherical symmetry \citep[e.g.][]{2008gady.book.....B,2008MNRAS.390.1647B}:

\begin{eqnarray} 
M_{\rm 3D}(r) & = & M(r) - 4r^2 \int_0^{\frac{\pi}{2}} \Sigma\left(x\right) \left(\frac{1}{\cth^2} \right. \nonumber \\ 
& & \left. - \frac{\sth}{\cth^3} \arctan\left(\frac{\cth}{\sth}\right) \right) \dth
\end{eqnarray}
where $M(r)$ is the projected enclosed mass evaluated at 3D radius $r$; and $x = r/\cth$.

This de-projection algorithm was tested on triaxial figures in
\citet{2006ApJ...652L...5S}. They found that for triaxialities typical of our
current cosmology, the method works extremely well unless the triaxial figure
is projected directly along the line of sight such that we see the galaxy or
galaxy cluster `down the barrel'. Such a situation is unlikely, but in any case
avoidable since the resultant figure appears spherical in projection. This
leads to the seemingly counter-intuitive result that the kinematic constraints
-- that rely on the above de-projection -- are most secure for systems that do
not appear spherical in projection (unless independent data can confirm the
three dimensional shape is indeed very round).

We use the deprojected mass to numerically solve equation
\ref{eqn:sphericaljeans} for constant $\beta(r)$, assuming either $\beta(r) =
1$ or $\beta(r) = 0$ at all radii to bracket the two extremum situations. Where
the data are good enough, these two may be distinguished giving dynamical
information about $\beta(r)$. In more typical situations, however, we seek to
simply marginalise over the effect of $\beta(r)$, using the stellar kinematics
as a robust measure of $M_{\rm 3D}(r_{1/2})$ (see \S\ref{sec:kinematics}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The mock data}\label{sec:mockdata}

We now present a study of four mock galaxies with known analytic forms. These
are used to verify that \Glass\ is able to correctly recover the mass profile,
and -- more importantly -- to determine what type and quality of data best
constrain the mass profile and shape of a lens.

\subsection{The triaxial N-body mock galaxies}

We generate four two-component mock galaxies, where the dark matter and stellar
profiles are allowed to be both steep and shallow.  
%The enclosed mass in all cases
%is fixed at $M(a_\star)= 1.8\e{10}$\,M$_\odot$ at the scale radius $a_\star=2$\,kpc. 
The enclosed mass of the stars and dark matter are both fixed to be 
$M_{*,\mathrm{DM}} = 1.8 \times 10^{10}$\,M$_\odot$ at the stellar scale radius 
$a_* = 2$\,kpc, such that the stars and dark matter contribute equally to the total
mass at $a_*$. The dark matter scale length is fixed for all models at $a_\mathrm{DM}
= 20$\,kpc.
These
values were chosen to closely resemble the lensing galaxy B1115 \citep{1980Natur.285..641W}. We place the galaxy at
a redshift of $z_L = 0.31$ for lensing.  Throughout, we assume a cosmology
where $H_0^{-1}=13.7$ Gyr, $\Omega_M=0.28$, and $\Omega_\Lambda=0.72$. The
critical density is $\kappa_\mathrm{crit}\sim 1.8\e{9}$\Msun/kpc$^2$.

The galaxies were generated as three dimensional particle distributions as in
\citet{2009MNRAS.395.1079D}. Each component follows the profile:
%
\begin{equation} 
  \rho(m) = \frac{M}{4\pi a^3}(3-\gamma){(m/a)^{-\gamma}(1 + m/a)^{\gamma-4}} 
  \label{Dehnen profile} 
\end{equation} 
%
where $a$ is the component scale radius mentioned in \tabref{mock galaxy params};
$m^2 = (x/\lambda_1)^2 + (y/\lambda_2)^2 + (z/\lambda_3)^2$ is the ellipsoidal radius; and the axis
ratios are $\lambda_1:\lambda_2:\lambda_3 = 6:4:3$.
In the case where the central density profile index
$\gamma$ is unity (and in the limit of spherical symmetry), this is the Hernquist
profile \citep{1990ApJ...356..359H}.  The four combinations of profile indices
are shown in \tabref{mock galaxy params}.

\begin{table}
\begin{tabular}{llllllll}
Galaxy & $\gamma_\star$ & $M_\star$ & $\gamma_\mathrm{DM}$ & $M_\mathrm{DM}$ & $\Rmap$ \\
\hline
\mockAA & 1 & 4 & 0.05 & $11^{2.95}$ & 50 kpc\\ % AA
\mockAC & 1 & 4 & 1 & $11^2$ & 50 kpc \\ % AC
\mockBB & 1.5 & $2^{1.5}$ & 0.16 & $11^{2.84}$ & 50 kpc \\ % BB
\mockBC & 1.5 & $2^{1.5}$ & 1 & $11^2$ & 10 kpc %BC
\end{tabular}
\caption{Profile parameters for the four mock galaxies. The name indicates whether 
  the galaxy is centrally dark matter or stellar dominated with a shallow or cuspy 
  dark matter density profile.  Masses are in units of
  $1.8\e{10}\Msun$. The scale lengths for all lenses are
  $(a_\star,a_\mathrm{DM})=(2,20)$\,kpc. $\Rmap$ is the 2D projected radius
used to generate the lens configurations. In the case of \mockBC, the profile is 
sufficiently steep that the profile could be truncated at $\Rmap = 10$ kpc.}
\label{mock galaxy params}
\end{table}

\begin{figure*}
\includegraphics[width=0.33\textwidth]{MockGalProfile-a.pdf} 
\includegraphics[width=0.33\textwidth]{MockGalProfile-b.pdf} 
\includegraphics[width=0.33\textwidth]{MockGalProfile-c.pdf}
\caption{
Profiles of the four mock galaxies showing the stellar (dotted) and dark matter (dashed) components and the total (solid).
\textbf{Left:} 
The spherically averaged density. The stars in models \mockBB{} and \mockBC{} contribute significantly to the central potential. 
\textbf{Middle:} 
The radially averaged two-dimensional projected density.
The critical density for lensing at $z_L=0.31$, $\kappa_\mathrm{crit}\sim 1.8\e{9}$\Msun/kpc$^2$, is marked by the horizontal line. 
\textbf{Right:}
The enclosed projected mass.
}
\label{mock galaxies}
\end{figure*}

In \figref{mock galaxies}, we show the 3D radial density, 
the 2D projected density, and the 2D enclosed mass for each
galaxy.

\subsection{Lens configurations}\label{sec:lensconfig}

For each of the four galaxies, we used the raytracing feature of \Glass\
described in \secref{Raytracing} to construct 6 basic lensing morphologies:

\begin{enumerate}
\item one double and one extended double;
\item one quad and one extended quad;
\item two 2-source quads with varying redshift contrast.
\end{enumerate}
The `extended' configurations use multiple point sources at the same redshift
to simulate an extended source that will produces an arc-like image.
\figref{arrival surfaces} shows the lens configurations for the \mockBC\
galaxy.  The configurations for the other galaxies are similar. Within the
names, the labels R1, R2, R3, refer to the redshift of the sources. We have
chosen R1=1.72, R2=0.72, and R3=0.51 so that the radial distribution of the
images is nearly equal.

\begin{figure*}
\includegraphics[width=0.85\textwidth]{BCarrival_surfaces}
\caption{The lens configurations for the six test cases using the \mockBC{}
mock galaxy. The other mock galaxies produce similar results. Here, the central
image is shown, although not all tests include it. The naming convention
indicates the redshift of the sources with R1=1.72, R2=0.72, and R3=0.51.  The
central image belongs to only to the R1 source to avoid overconstraining the
models. Image colors represent a common source. Grey circles are a visual aid to help determine radial
separation between images. The axes are in arcseconds.}
\label{arrival surfaces}
\end{figure*}

Each of these configurations were modelled with and without time delays; with
and without a central image; and with and without the stellar mass as a lower
bound, for a total of 48 test cases. (The central image is typically highly
demagnified. For galaxy lenses it is very difficult to find since it lies along
the sight line to the bright lensing galaxy; in clusters, however, such images
have been seen -- e.g. \citealt{2005PASJ...57L...7I}). We assumed for all our tests that the
lensing mass was radially symmetric (Prior vi). For our mock data, this is
known to be true; it is most often the case with real galaxies, unless there is
an obvious observed asymmetry. The central pixel was refined into a further
$5\times5$ pixels (three extra radial positions) to capture any steep rise in
the profile.  Two of the four mock galaxies have a steeply rising inner
profile. \hilight{JR.  CAN WE DO THIS? What happens if we don't assume this symmetry? We ought
to add an appendix showing this for one example case.}

In all cases, we additionally assume no external shear. This amounts to an assumption that any such shear field can be constrained by data (e.g. combining weak lensing constraints, or assuming that the shear field correlates with visible galaxies -- e.g. \citealt{2009A&A...500..681M,2011ApJ...726...84W}).

\section{Results}\label{sec:results}

\subsection{Radial profile recovery}

\figref{reconstruction} shows some example reconstructions of the radial profile of our mock lenses. The left
column shows the ensemble average arrival time surface with images marked as
circles and the inferred source position(s) as a diamond. The centre column shows
the radial density profile. The error bars cover a $1\sigma$ range around the
median; the grey bands show the full ensemble range. The true density profile from the mock data is also plotted for
comparison. The vertical lines mark the radial position of the images. The
right column shows the enclosed mass. From top to bottom, the rows correspond to an extended double for \mockBC; an extended double with stellar mass constraints for \mockBC; a quad with time delay data for \mockBC; and a quad with time delays for \mockAA. \figref{2d mass reconstruction} shows an example 2D reconstruction for \mockBC\ for a quad; we discuss shape recovery further in \S\ref{sec:shape}.

As expected, the error bars
are smallest in the region of the images where the most information about the
lens is present. In all cases, there is a dip in the profile at large $R$ due to the
cut off in mass in the lensing map. This is of little importance, though, as
there is no lensing information there.

Notice that the extended double (top row) gives the poorest constraints, as expected. Adding stellar mass (second row) significantly improves the constraints, for this example where the stars contribute significantly to the potential. Moving to a quad with time delays gives constraints almost as strong as the double with stellar mass, but note that {\it focussing only on the goodness of the fit can be misleading.} In the third row of \figref{reconstruction}, we obtain a better recovery than in the bottom row for {\it precisely the same data quality}. This occurs because the \Glass\ prior favours steeper models consistent with \mockBC, but not \mockAA. It is the \Glass\ prior, rather than the data that is driving the good recovery for \mockBC\ in this example. This emphasises the importance of using a wide range of mock data tests to determine the role of data versus prior in strong lensing.

\begin{figure*}
  \plotthree{BCExtendedDoubleR1_tms-a.pdf} {BCExtendedDoubleR1_tms-b.pdf} {BCExtendedDoubleR1_tms-c.pdf}
  \plotthree{BCExtendedDoubleR1_tmS-a-1.pdf} {BCExtendedDoubleR1_tmS-b-1.pdf} {BCExtendedDoubleR1_tmS-c-1.pdf}
  \plotthree{BCQuadR1a_Tms-a.pdf} {BCQuadR1a_Tms-b.pdf} {BCQuadR1a_Tms-c.pdf}
  \plotthree{AAQuadR1a_Tms-a.pdf} {AAQuadR1a_Tms-b.pdf} {AAQuadR1a_Tms-c.pdf}
\caption{
    Four reconstructed lenses. The first two rows consider the same mock
    galaxy, \mockBC, for an extended double, and highlight the improved
    constraints on the mass distribution when a lower bound is given by the
    stellar mass. For the reconstruction, no time delays were assumed.  The
    third row uses again the the \mockBC{} mock galaxy, but with time delays
    for a single quad and no stellar mass. With the added information from the
    quad, the outer regions of the lens are the image radii are better
    constrained. Finally, the fourth row shows another case of a quad with time
    delays, but using the \mockAA{} mock galaxy. This galaxy has a shallower
    stellar density index, and a core in the dark matter. Due to the priors
    used in \Glass, the modelling favours steep solutions without additional
    information.
\textbf{Left:}
The ensemble average arrival time surface with just the iso-contours for the
saddle points drawn. The central diamonds show the reconstructed source
position.
\textbf{Middle:}
The surface density of the dark matter (DM; magenta); the stars (yellow); and the total (blue).
The original mass model used to create the lens is shown in black.
The vertical lines mark the radial positions of the images. The higher
resolution feature of \Glass\ has been used on the central pixel allowing the
steep profile to be captured.
\textbf{Right:}
The cumulative mass. The error bars on all plots are $1\sigma$; the grey bands show the full range of models.}
\label{reconstruction}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS-kappa-a.pdf}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS-kappa-b.pdf}
\caption{ \textbf{Left:} The mock data distribution for \mockBC{} projected onto a coarse grid.
  \textbf{Right:} The recovered ensemble average $\kappa$ distribution for the single quad. The contours are
  logarithmic base 10 values, where level 0 corresponds to the critical density. Contours below the critical density
  are drawn with dashed lines.}
\label{2d mass reconstruction}
\end{figure*}

\figref{main results} and \figref{main results pixel-wise} show the results for our full mock data ensemble. Each
subplot corresponds to a different mock galaxy, as marked. We show the fractional error of the mass
distribution for each of the test configurations with (red) and without (black)
 stellar mass. In \figref{main results}, the error is defined over radial bins of mass as:
%
\begin{equation} \label{ferror R}
  f_R = \frac {\sum_i \left|M(i) - \tilde M(i)\right| } {\sum \tilde M(i)}
\end{equation}
%
where $M$ is the mass in bin $i$ and $\tilde M$ is the mass from the mock galaxy.
In \figref{main results pixel-wise} the error is defined over all the pixels $\vec\theta$:
%
\begin{equation} \label{ferror theta}
f_\theta = \frac {\sum_{\vec\theta} \left|M(\vec\theta) - \tilde M(\vec\theta)\right| } {\sum_{\vec\theta} \tilde M(\vec\theta)}
\end{equation}
With both $f_R$ and $f_\theta$, we only consider mass up to one radial
position passed the outermost image, since there is no longer any lensing
information beyond that point.

\begin{figure*}
\includegraphics[width=0.49\textwidth]{AAferror_profile-1sig.pdf}
\includegraphics[width=0.49\textwidth]{BBferror_profile-1sig.pdf}\\
\includegraphics[width=0.49\textwidth]{ACferror_profile-1sig.pdf}
\includegraphics[width=0.49\textwidth]{BCferror_profile-1sig.pdf}
\caption{Our main results showing the quality of the radially averaged model recovery \eqnrefp{ferror R} for all our test cases. 
Within each panel are six groups of results for each of six lens morphologies. Each morphology
considered the presence of time delays (T=yes,t=no) and a central image (M=yes,m=no). The black markers are for tests
that did not include the stellar mass as a lower bound constraint, while the red markers
indicate where the stellar mass has been included. Error bars show the $1\sigma$ range of the model ensemble.}
\label{main results}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.49\textwidth]{AAferror-1sig.pdf}
\includegraphics[width=0.49\textwidth]{BBferror-1sig.pdf}\\
\includegraphics[width=0.49\textwidth]{ACferror-1sig.pdf}
\includegraphics[width=0.49\textwidth]{BCferror-1sig.pdf}
\caption{Similar to \figref{main results} but for the fractional error
in the pixel-wise recovery \eqnrefp{ferror theta} of all models. The colours and labels are the same
as previously. Error bars show the $1\sigma$ range of the model ensemble.}
\label{main results pixel-wise}
\end{figure*}

The abundance of strong lensing data increases from left to right within each
plot. As a result, there is a general trend for the reconstruction quality to
increase (and therefore for $f$ to decrease). When both time delays and
a central image are present (TM), the quality is highest. A double is known to
provide very little constraint on the mass distribution. This is particularly
evident in galaxies \mockAC{} and \mockBC{} where the mass profile is steepest
and the reconstruction of the double is poorest. However, the addition of an
arc from the extended source is sufficient to correct this. Notice that, as in \figref{reconstruction}, the recovery for \mockBC\ quickly saturates; there is little improvement as the data improves beyond a single quad. This occurs because the \Glass\ sample prior in the absence of data favours steep models like \mockBC\ over shallower models like \mockAA\ (see also \figref{reconstruction}).

\subsection{Shape recovery}\label{sec:shape}

\figref{main results pixel-wise} already gives us important information about how well we can recover the {\it shape} of a lens. The trends are very similar to the radial profile recovery in \figref{main results}, suggesting that if the radial profile is well-recovered then, typically, the shape is too. A notable exception is for the \mockBC\ models where adding stellar mass constrains aids the shape recovery, but little-improves the radial mass profile. A visual example of the shape recovery is given in \figref{2d mass reconstruction}.

We can also more directly probe the recovery of the shape of the mass
distribution by defining a global measure of lens shape. We define this to be the ratio of the eigenvalues of the 2D inertia tensor:

\begin{equation} 
I_{ij} = \sum_\theta M(\vec\theta) \theta_i\theta_j \,\,\, ; \,\,\, s \equiv \lambda_1/\lambda_2
\label{eqn:inertia}
\end{equation}
where $\lambda_1$ is always taken to be the largest value. As with $f_R$ and $f_\theta$, we only consider mass up to one radial position passed the outermost image and compute the fractional error as:
%
\begin{equation} \label{ferror shape}
  f_\mathrm{shape} = \left|s - \tilde s\right|  / \tilde s
\end{equation}
%
where $\tilde s$ is the shape of the mock galaxy. The distribution of $f_\mathrm{shape}$
for each mock galaxy and each test case is shown in \figref{shape results}. Interestingly, for this global shape parameter recovery it appears more important to have time delay data and/or a central image (T/M) than to have a quad or multiple sources with wide redshift separation. In all cases, the stellar mass little-aids the recovery, reflecting the fact that $s$ is heavily weighted towards the shape at the {\it edge} of the mass map, rather than at the centre where the stars may dominate the potential (see equation \ref{eqn:inertia}).

\begin{figure*}
\includegraphics[width=0.49\textwidth]{AAferror_shape-1sig.pdf}
\includegraphics[width=0.49\textwidth]{BBferror_shape-1sig.pdf}\\
\includegraphics[width=0.49\textwidth]{ACferror_shape-1sig.pdf}
\includegraphics[width=0.49\textwidth]{BCferror_shape-1sig.pdf}
\caption{Here we demonstrate our ability to recover the shape of the lensing mass. The shape ratio
$\lambda_1/\lambda_2$ is measured from the principal components $\lambda_1, \lambda_2$ of the mass up to the outermost image.
We plot the distribution of fractional error compared with the shape of the mock galaxies \eqnref{ferror shape}.}
\label{shape results}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_Tms_sigp-1.pdf}
\includegraphics[width=0.33\textwidth]{BCQuadR1a_TmS_sigp-2.pdf}
\includegraphics[width=0.33\textwidth]{BC_beta.pdf}
\caption{ Estimated projected radially averaged velocity dispersion $\sigma_p$
\eqnrefp{eqn:sphericaljeans} for a single quad from the \mockBC{} mock galaxy without stellar mass
(\textbf{left}) and with stellar mass (\textbf{middle}) assuming an anisotropy $\beta=0$
(blue, solid) and $\beta=1$ (blue, dashed).  The equivalent
curves are also shown for the projected mock data after using the same analysis routines
(purple). The solid black line is the actual cylindrically averaged velocity
dispersion of the original mock particle data.  The stellar half mass radius
(yellow) Einstein radius (black) are marked by vertical lines. For this
configuration, these two radii are well-separated. The actual variation in $\beta(r)$ is also shown (\textbf{right}).}
\label{fig:sigp} \end{figure*}
\subsection{Stellar mass}
\label{stellar mass}

The stellar mass distribution gives a lower bound on the total mass. Where the
stars dominate the central potential, it can provide a powerful constraint
extra to the strong lensing data. We took the stellar mass directly from the
generated galaxies and projected the particles onto the pixels. \Glass\ also
offers an option to interpolate any map of stellar mass (e.g., from an
observation) onto the pixels. The linear constraint is added to \Glass\ by
writing $\kappa_n = \kappa_{dm,n} + \kappa_{s,n}$ as the sum of the dark matter
and stellar mass components in the potential \eqnrefp{discrete potential}.
Since each $\kappa_{s,n}$ is just a constant we do not add new, separate
equations for each pixel. Although we assume a perfect recovery of the stellar
mass here, it is straightforward to add errors as the stellar mass constraint
remains linear: $\kappa_n = \kappa_{dm,n} + \epsilon \kappa_{s,n}$, where
$\epsilon \sim 1$ is an additional error parameter.

With the stellar mass lower bound, there is a significant improvement of the reconstruction
quality shown in \figref{main results} and \figref{main results pixel-wise} for the doubles in the
steepest mock galaxies (\mockAC{} and \mockBC). This is because these models
are dominated by stars in the inner region. By contrast, the other two galaxies
 --- where the stars contribute negligibly to the potential --- are largely unaffected.

\subsection{Stellar kinematics}\label{sec:results_stellar_kinematics}

As outlined in \S\ref{sec:glass}, \Glass\ can also run post processing routines
on the model ensemble which can be used to apply non-linear constraints. As an
example, we consider here constraints from stellar kinematics. The models in
the \Glass\ ensemble are processed as described in \S\ref{sec:glasskinematics}.
To illustrate the power of stellar kinematic constraints, in
\figref{fig:sigp}, we plot the projected velocity dispersion calculated for
one model model (extracted from the full ensemble) of the \mockBC{} Quad with time
delays and stellar mass (left), and the same but without stellar mass (right).
In both cases, we calculate curves
for two extrema velocity anisotropies: $\beta=0$ (green) and $\beta=1$ (red).
Over-plotted is the correct answer for the \mockBC{} model (black). The stellar half
mass radius (yellow) Einstein radius (black) are marked by vertical lines. For
this configuration, these two radii are well-separated.

Without even sweeping through the model ensemble and formally
accepting/rejecting models, \figref{fig:sigp} already illustrates what we can obtain from stellar kinematics. The left plot shows the radially averaged projected velocity dispersion $\sigma_p(R)$ \eqnrefp{eqn:sphericaljeans} for a single quad from the \mockBC\ galaxy without the stellar mass constraint. The blue data points show the 1$\sigma$ distribution from the ensemble assuming $\beta=0$ (solid) and $\beta = 1$ (dashed); the grey bands show the full distributions. Also marked are the $\sigma_p(R)$ calculated from the mock data assuming $\beta=0$ (solid purple) and $\beta =1$ (dashed purple); and the true $\sigma_p(R)$ measured directly from the stars (black). This latter has a non-constant $\beta(r)$ (right panel) and differs also from the purple and blue curves in that these all assume spherical symmetry, whereas the stellar distribution is really triaxial. Such triaxiality and varying $\beta(r)$ explains why the purple curves do not match the black one. However, they do largely bracket the correct solution. More interestingly, the curves approximately cross for $\beta = 0$ at the stellar half light radius (yellow vertical line). This demonstrates, as has previously been reported in the literature, that $\sigma_p(R)$ gives a good estimate of the mass enclosed within $\sim$ the half light radius $M_{1/2}$ \citep[e.g.][]{2009ApJ...704.1274W,2010MNRAS.406.1220W}. The mass {\it profile}, however, depends on $\beta$ which is poorly constrained by these data. If we add stellar mass constraints (middle panel), the situation is little-improved. The true answer already lay close to the bottom of the ensemble distribution; it now is forced to lie right at the edge.

From \figref{fig:sigp}, it is clear that $\sigma_p(R)$ provides two useful pieces of information. Firstly, it is a powerful probe of $M_{1/2}$. Given a measurement of $\sigma_p(r_1/2) \sim 150$\,km/s, we could usefully reject many models in the ensemble as being overly steep in the centre. We would not, however, obtain a strong constraint on $\beta(r_{1/2})$. We could rule out $\beta(r_{1/2}) = 1$ (blue dashed line), but since our $\beta=0$ model crosses the true $\beta \sim 0.5$ line at $r_{1/2}$ it is clear that many $\beta(r)$ profiles will be consistent with the data. On the other hand, if we have a situation where $r_{1/2} \sim r_E$ (i.e. the vertical yellow and black lines in \figref{fig:sigp} overlap), then we will obtain tight constraints on $\beta$ since we then have two strong constraints on $M(r_{1/2})$ that become redundant. This latter situation of redundancy is also exactly what we would like to constrain cosmological parameters. In this case, we require a third piece of redundant information -- in this case in the form of strong lensing time delays. We will discuss such cosmological constraints in a forthcoming paper.

The results for stellar kinematics match our expectations from
\S\ref{sec:kinematics}. Where the lens data already constrain the mass
distribution at $r \sim r_{1/2}$, stellar kinematics provide valuable
information about the velocity anisotropy of the stars, $\beta$ (see
\figref{fig:sigp}). Where the lens data poorly constrain the mass distribution
at $r_{1/2}$, we may `integrate out' the effect of unknown $\beta$ to obtain a
robust measure of $M_{\rm 3D}(r_{1/2})$ from the stellar kinematics. This latter is
robust to both uncertainties in $\beta(r)$ and to our assumption of spherical
symmetry in the kinematic models \citep{2012ApJ...754L..39A}.

\section{Conclusions}\label{sec:conclusions}

We have introduced a new gravitational lens modelling tool ---\Glass--- and used
it to test the recovery of the mass profile and shape of mock strong lensing
galaxies. Our key findings are as follows:

\begin{enumerate}
\item For pure lens data, multiple sources with wide redshift separation give the strongest
constraints as this breaks the well-known mass-sheet or steepness degeneracy;

\item A single quad with time delays also performs well, giving a good recovery
of both the mass profile and its shape;

\item Stellar masses -- for lenses
where the stars dominate the central potential -- can also break the steepness
degeneracy, giving a recovery almost as good as having time delay data or
multiple source redshifts;

\item If the radial density profile is well-recovered, so too is the shape of a lens;

\item Stellar kinematics provide a robust measure of the mass at the half light radius of the stars $M(r_{1/2})$ that can also break the steepness degeneracy if $r_{1/2} \neq r_E$ -- the Einstein radius; and

\item If $r_E \sim r_{1/2}$, then stellar kinematic data can be used to probe the stellar velocity anisotropy 
$\beta$ -- an interesting quantity in its own right.

\end{enumerate}

Where information on the mass distribution from lensing and/or other probes becomes
redundant, this opens up the possibility of using strong lensing to constrain
cosmological models. We will study this, and present the first results from
\Glass\ applied to real data, in forthcoming papers.

\section{Acknowledgments}\label{sec:Acknowledgements}

The authors would like to thank Sarah Bryan and Walter Dehnen for creating the
particle distributions for the mock galaxies. JIR would like to acknowledge support from SNF grant PP00P2\_128540/1.

\appendix

\section{Implementation Details}

Since we want to model the density distribution with a computer it
is convenient to choose units that make the relevant quantities of order
unity.  We therefore measure lengths in light years, time in years, positions
in arcseconds, and choose $c=1$ and $4\pi G = N^2$, where $N^2 \equiv 206,265$
arcsec/rad. The mass unit is then $11.988\ \Msun$. It will also be useful to
define a proxy to the Hubble constant $\zeta \equiv N^2 H_0$. We now 
express the equations from \secref{sec:theory} in terms of these new units
and introduce some other useful quantities.

The lens equation in its complete form becomes:
%
\begin{eqnarray}
N^2ct(\vec\theta) & = & (1+z_L)\frac{D_{L}D_{S}}{D_{LS}}\frac12 |\vec\theta - \vec\beta|^2 \nonumber \\
& & - (1+z_L)\frac{4GD_{L}^2}{c^2}\int \Sigma(\vec\theta') \ln |\vec\theta-\vec\theta'| d^2\vec\theta'
\label{eqn:full_arrival_time}
\end{eqnarray}
%
where the factor of $D_L^2$ in the second term comes from the fact that $\Sigma$
has units of \Msun/lyr$^2$. We can clean this up by first writing down a dimensionless time
delay
%
\begin{equation}
\tau = \left[ (1+z_L) d_L\right]^{-1}\zeta t
\label{tau}
\end{equation}
%
in terms of our previous definitions and defining $D_L \equiv (c/H_0)d_L$. We
further define a dimensionless density
%
\begin{equation}
\kappa_\infty = \frac{4\pi G}{c^2}\frac{c}{H_0}d_L\Sigma
              = \frac{d_L}{\zeta}\Sigma
\end{equation}
%
and a lensing potential
%
\begin{equation}
\psi(\vec\theta) = \frac1\pi \int \kappa_\infty(\vec\theta') \ln|\vec\theta - \vec\theta'| d^2\vec\theta'\
\label{lensing potential}
\end{equation}
%
Now we can express \eqnref{eqn:full_arrival_time} very compactly as
%
\begin{equation}
\tau(\vec\theta) = \frac12 \xi |\vec\theta-\vec\beta|^2 - \psi(\vec\theta)
\label{arrival time}
\end{equation}
%
where $\xi=d_S / d_{LS}$.
We explicitly write $\kappa_\infty$ to remind ourselves
that there is no source distance factor involved. This is useful when we
consider multiple sources.

%\section{Raytracing convergence test}
%Raytracing is sensitive to the map area and resolution set by $\Rmap$ and
%$\Rpix$.  In the left panel of \figref{raytracing convergence tests} we compare
%the change in time delays as we increase $\Rmap$.  The image system is a quad
%with the central maximum image. When the projected mass begins to fall off the
%time delays stabilize. In the right panel we see the effect of changing the
%resolution of the map.  Increasing the resolution allows for more accurate
%placement of the images and after about $\Rpix=40$ the image positions change
%by less than 0.04 arcsec. The one image that continues to change is near to a
%caustic.  
%
%\begin{figure}
%\plottwo{tdconv_pr45.pdf}{imgpos_conv_mr20.pdf}
%\caption{\textbf{Left:} Test for convergence of the predicted time delay as $\Rmap$ is increased and
%$\Rpix=45$ is held constant. After about $\Rmap=20$ most of the asymmetric mass is in the map.
%\textbf{Right:} Test for convergence of the predicted image positions $\theta$ as $\Rpix$
%changes and $\Rmap=20$ arcsec is held constant.}
%\label{raytracing convergence tests}
%\end{figure}

\section{Derivation of pixelated density coefficients}
\label{Q derivation}
When the lens plane is pixelized we need a discrete form of the integral
%
\[\int \kappa(\vec\theta') \ln |\vec\theta-\vec\theta'| d^2\vec\theta' \]
%
In particular we want
%
\[\sum_n \kappa_n Q_n(\vec\theta)\]
%
where $Q_n$ is the logarithm evaluated over the $n$th pixel at position $\vec\theta_n = (x_n, y_n)$. Let the pixel side length be $a$.
Instead of working with a position vector $\vec\theta$ we work in cartesian coordinates where 
%
$|\vec\theta| = r = \sqrt{x^2 + y^2}$. The integral now becomes
%
\[Q_n(x,y) = \frac12 \int_{y_-}^{y_+}\int_{x_-}^{x_+} \ln (x'^2+y'^2) dx' dy'\]
%
where $x_\pm = x + x_n \pm (a/2)$ and similarly for $y_\pm$.
Using the identity
%
\[\int \ln(x^2+y^2) dx = x \ln(x^2+y^2) - 2x + 2y\arctan(x/a) \]
%
we can express $Q_n$ as the sum of four parts
%
\begin{align*}
Q_n(x,y) = \frac12 [\tilde Q_n(x_+,y_+)
                  + \tilde Q_n(x_-,y_-) & 
\\                - \tilde Q_n(x_-,y_+)
                  - \tilde Q_n(x_+,y_-) ]
\end{align*}
%
where
%
\[\tilde Q_n(x,y) = xy(\ln r^2 - 3) + x^2\arctan(y/x) + y^2\arctan(x/y)\]

\section{Pixel resolution convergence test}
\hilight{JR. Need to add an appendix showing convergence with increasing Pixrad.}

\bibliographystyle{mn2e}
\bibliography{ms}

\end{document}

\cite{2007ApJ...667..645R} were the first to test the recovery of
strong lensing surface density profiles using mock data generated from
N-body models {\bf JR. Is this true?}. They found that when including
time delays, circularly averaged surface density distributions can
sufficiently well-recovered to make interesting comparisons with
theoretical models Cosmological parameters, however, are more
sensitive. Since then, several groups have used N-body mock data to
explore the fidelity of strong lensing reconstructions
\citep{2007MNRAS.380.1729L}. We expand on these earlier works
by considering not only the recovery of the circularly averaged
surface density profile of a lens, but also its shape, higher order
measures of the mass recovery (a pixel-by-pixel comparison), and the
utility of wrapping in constraints from stellar mass estimates and/or
stellar kinematics.