### 2015-09-29 shapes/reply_referee_1.txt
Final changes

We would like to thank the referee for a constructive and helpful report that has improved our paper. We have now addressed all of the referee's previous concerns and we hope that the paper is now ready for publication. We respond to each point in turn, next.

| Examples of lack of clarity:

| - The authors state that "isolated misaligned galaxies are expected to
|   be unstable" but they do not cite any references. What do the authors
|   mean by misalignment (e.g., mass or angular momentum)? Debattista et al.
|   (2013) showed that disc galaxies with their spins along the intermediate
|   axis aren't stable, but they also showed that when gas is present a disc
|   can survive for a very long time if it is off of one of the symmetry
|   planes of the halo. Debattista et al. (2015, arXiv:1502.03429) simulated
|   disc formation with gas and found that discs can have essentially arbitrary
|   orientations. There is some observational evidence from the observed
|   locations of satellite galaxies (e.g., Agustsson & Brainerd 2010) and the
|   intrinsic alignments of galaxies (e.g., Singh et al. 2014, arXiv:1411.1755
|   and references therein) that dark matter mass and light may not be well
|   aligned in disc galaxies (i.e., the intrinsic alignment signal is most
|   robustly detected for red galaxies, not blue galaxies).

We have now clarified this in a new section (5.2) on the lack of misalignment in isolated lenses. We more carefully define "isolated" as having external shear $\gamma < 0.1$, and we now discuss the fact that gas rich galaxies can maintain misalignments over long times, while gas poor galaxies cannot. Our lenses are all early types except 2237 for which we probe only the bulge region. Thus we expect them to be gas poor "red" galaxies that are, in the absence of external shear, aligned.

| - The authors say that they assume "the local density gradient to be
|   always pointing within 50 degree of the centre of the
|   distribution". What does this mean? They also say that "the radius
|   of the lens is set to be 12 pixels". What does this mean? What is
|   the radius of the lens? (The optical radius perhaps?) Presumably
|   the pixels refer to what is used in the lens reconstruction but
|   this is not stated sufficiently clearly.

Section 2 (which describes the method) has now been rewritten and is hopefully clearer. The unclear part about the local density gradient has been replaced with detailed references. On the mass-pixel size, we now explain that they are at most 4% of the diameter of the full mass map, while the stellar/dark mass map extends to the outermost image. More explicit details of the methodology are covered in Coles et al. 2014.

| - The authors do not specify the physical radii over which they are
|   attempting to constrain the dark matter mass. How does the scale being
|   probed compare to, say, the virial radius that might be expected?

We have included an additional Figure (Figure 1), which besides addressing other questions as well, illustrates the range in physical scales probed for every lens system. Furthermore, as we elaborate below, we have calibrated NFW profiles to match the convergence profile of the reconstructed distribution at the Einstein radius. The virial masses and the corresponding virial radii derived in this way are now given in Table 3.

| - It is not clear what form the best-fitting gravitational potentials of
|   the dark matter halos turn out to be. Are they consistent with some
|   reasonable expectation for galaxy formation (say, elliptical NFW or
|   elliptical isothermal), or are they something completely different to
|   naive theoretical expectations?

We planned to present the radial dark matter profile of these lenses as part of a larger sample in a separate paper. However, we agree with the referee that it is useful to present these here also. We have addressed this issue by including several new Sections and Figures: Firstly, Section 4.1 was added. It compares the projected dark matter profile of each individual lens with an NFW profile, assuming the cosmic mean concentration parameter. Secondly, Figure 1 has been included displaying the reconstructed azimuthally averaged convergence and the NFW profiles as a comparison. We derive virial masses and radii by matching NFW profiles to the measured projected dark matter mass profiles at the outermost image. The values derived in this way are given in Table 3. As can be seen, our lenses are all well fit by an NFW profile or are steeper, as expected if halos contract in response to gas cooling. We have added these new results also to the abstract and conclusions. An interesting new result that arose from this is that those lenses with an outermost image inside ~10kpc appear to be contracted with respect to NFW; while those with an outermost image >10kpc are rather well fit by the NFW form. We discuss potential reasons for this, but defer further analysis to a separate publication (where we will vary, for example, the stellar IMF).

| - It is not clear how the best-fitting values for the halo shape and
|   degree of alignment with the stellar light are determined. Similarly
|   it is not clear how the error bars on these values (and their significance)
|   were determined.

We have reworked Section 2 significantly to explicitly describe our estimators and the statistical uncertainties for the shape and orientation of the distribution. These quantities are estimated, as now explained in the text, by computing the eigenvalues and eigenvectors of the inertia tensor (see Eq. 1). Shape and orientation (resp. misalignment) are then estimated by combining the two eigenvalues resp. the angle of the eigenvector corresponding to the larger eigenvalue relative to the x-axis (see Eq. 2 and 3). As Glass gives an ensemble of solutions, these quantities are estimated for each solution. Based on these distributions, the quantities are constrained and the statistical uncertainties estimated.

| - Regarding 0957, the authors say "the position angles to the center of
|   the cluster of earlier lens constructions ... " which is too vague. What
|   the authors mean is the position angle of the external shear. Throughout,
|   the authors need to specify "the position angle of _______" (i.e., by
|   explicitly filling in the blank) so that readers can understand what
|   they're trying to say.

We have made this clearer along the lines suggested by the referee. We now add these angles also to Table 3 and more carefully explain that this position angle can be directly compared to the external shear angle.

| - Figure 1 attempts to compile all of the main results into one figure, and
|   it is extraordinarily difficult to disentangle this information. The bulk
|   of the data points lie on top of each other and it's not even possible to
|   tell which point and error bar corresponds to which lens system. At a minimum,
|   the authors need to include an additional table that lists, for each lens, the
|   values and error bars for e*, edm, the offset between the major axis of the
|   light and the major axis of the dark matter halo, and the external shear.

We thank the referee for the suggestions on how to avoid this slight over-compression of information in this key Figure. We have made several changes to improve this. First, we have added a second panel to this Figure, zooming into the region where the most points lie. Second, by changing the color scheme, understanding this third dimension in this Figure (external shear) becomes easier. Third, we provide the reader, as suggested, with all the values (Table 3). The ellipticities, for which we give the statistical uncertainties (see above), are now given using the ellipticity definition e=(a-b)/(a+b), which is more common in weak lensing. By changing the employed ellipticity definition, the comparison with the weak lensing results is facilitated (see below).

| - The authors do not make clear (with a statement of a value for γ)
|   for what they consider to be a "weak" external shear. To me, a
|   shear of more than a few percent is a relatively large external
|   shear, suggesting the presence of a larger mass concentration
|   within which the primary lens is embedded. That is, it seems that
|   even where no cluster/group members are currently known, there is
|   an external shear that could be explained by the presence of a
|   group/cluster environment (if I'm reading the colors in Figure 1
|   correctly).

The corresponding Sections have been adjusted and we now state clearly (see e.g. Abstract) that an external shear larger than >~0.1 is considered strong. Thus, the lens sample splits into two subsets, i.e. 6 lenses with a strong external shear, 5 with weak and intermediate external shear (see Table 3). For the lenses with a known and constrained environment, the direction of external shear (see Table 3) is consistent with the centre of the group/cluster (see Table 2). We agree with the referee that even an external shear of gamma ~ 0.08 might be considered "strong". However, we find that our lenses show qualitatively different behaviour for gamma < 0.1 and gamma > 0.1 and hence set the dividing line at gamma ~ 0.1.

| - The authors do not address how the direction of the external shear
|   compares to the orientation of the dark matter halo of the lens
|   galaxy, nor is there a full discussion how the magnitude of the
|   external shear compares with the inferred shape of the dark matter
|   halo.

Appendix B has been added to discuss the degeneracy between DM ellipticity and external-shear. Figure B1 now plots the scatter (over the model ensemble of each lens) of the orientations of DM vs external-shear. As can be seen, while there is some degeneracy between shape and shear, the degeneracy is largely broken and we are able to obtain good estimates of both. This finding and the dependence of the shape estimation on the shear prior are discussed explicitly in Section 4.2.

| Misinterpretation of previously published work:

| - It’s not correct that weak lensing can directly probe the orientation of the
|   dark matter mass relative to the luminous lens galaxies. The only way in which
|   one can even search for an anisotropy in the shear profile is to use the
|   symmetry axes of the luminous lens galaxies because you don’t know the
|   orientation of the symmetry axes of the dark matter. That is, the only sensible
|   way to stack the lenses is to use the major axes of the luminous galaxies. If
|   the signal you observe is isotropic, this could be telling you that the halos
|   are round or it could be telling you that mass and light are not well aligned
|   in your lens galaxies. You can’t disentangle those two possibilities from the
|   observations alone.

| - It’s not correct that Hoekstra et al. (2004) found “some weak evidence for
|   alignment”. Hoekstra et al. (2004) concluded that there must be quite good
|   alignment of mass and light for them to have detected a signal at all (for the
|   reason stated above). This is stated in their abstract.

| - It’s not correct that Mandelbaum et al. (2006) found that the halos of red
|   galaxies were flatter than their stellar distributions. They found the halos to
|   be rounder than the luminous galaxies in all cases (power law and NFW mass models).
|   This is stated in their abstract.

| - Given that Parker et al. (2007) do not quote error bars on their estimates of the
|   halo ellipticity, it’s not correct to say that they concluded the halos of red
|   galaxies were flatter than their stellar distributions. All of their halo
|   ellipticities are quoted with a “tilde” out front (e.g., ~0.3 and ~0.4). If a
|   proper error analysis were done, it could easily be found that these approximate
|   central values were consistent with the ellipticities of the stellar distributions.
|   Parker et al. (2007) put error bars on the shear anisotropy only and, properly, their
|   “2σ” statement in the abstract refers to their constraint on the shear anisotropy,
|   not the inferred halo ellipticity. In the full text they make it clear that they find
|   a 2σ detection of “non-sphericity”, which is not the same as a 2σ measurement of a
|   value of the projected ellipticity.

We thank the referee for these good points of criticism. We have revised the whole paragraph to give a more adequate description of the difficulties of this measurement and the results on data. We now describe what exactly can be constrained, which is, as pointed out by the referee, the projected dark matter ellipticity and thus cannot be directly interpreted as the ellipticity of the dark matter halo. Furthermore, the most recent observational constraints including the papers submitted in the meantime by Clampitt & Jain, 2015 and Schrabback et al., 2015 are presented.

| Additional Comments

| - The authors have not provided a sufficient description of GLASS
|   and its reliability for the reader. As it stands, the reader needs
|   to go to Coles et al. (2014) for too much relevant
|   information. Manuscripts need to be sufficiently independent of
|   one another that an “educated” reader need not be intimately
|   familiar with a preceding manuscript in order to understand and
|   evaluate the present manuscript.

As noted above, this Section has been rewritten to be more self-contained. It still refers the reader to Coles et al (2014) for precise formulations and tests, but without making it essential to consult the earlier work (unless interested in these explicit details and test results).

| - Nowhere in this manuscript do the authors make a compelling case
|   (through, e.g., model testing) that GLASS yields reliable results
|   for the quantities they seek.  Perhaps that has been done in
|   another manuscript but, if so, such tests have not been referenced
|   here explicitly. Because most of the lenses are clearly not
|   isolated, there is a concern about how the inferred external shear
|   may correlate with the inferred halo shape and orientation. The
|   authors need to demonstrate with reasonable theoretical models
|   that the magnitude and orientation of external shears in the range
|   of, say, 0.0 to 0.25 do not significantly affect the conclusions
|   regarding the shape and orientation of the dark matter halo of the
|   lens galaxy.

The revised Section 2 refers to Section 5.2 and Figure 8 of Coles et al (2014) which tests recovery of the shape estimators.

As noted above, Appendix B has been added to discuss relation between the orientation of the dark matter vs the direction of external shear.

| - In the concluding section, the authors restate their results but do not discuss them
|   in context with previous, similar work and known theoretical expectations. This is
|   important, and needs to be done – even when different techniques are used in the
|   observational analyses.

We agree that the paper can be improved in comparing our results to results using different techniques. To satisfy this, we have included a Discussion Section (Section 5). In this Section, we elaborate on the constraints from weak lensing. Especially, we compare our results explicitly to the study by Schrabback et al., 2015, submitted to MNRAS a short time ago, which is the most recent weak lensing measurement of halo ellipticity/misalignment, and furthermore studies the expected signal using the Milennium simulations. We facilitate this comparison furthermore by consistently using the same ellipticity definition as in Schrabback et al (see above), and computing the combination of ellipticities weak lensing can probe, i.e. f_h = e_dm / e_s, where e_dm is the dark ellipticity in direction of the light distribution and e_s is the ellipticity of the latter.

As we elaborate in the first Subsection 5.1, our constraints are broadly consistent with the most recent measurement on CFHTLenS data and the predictions from the Milennium simulation.

In a second Subsection, we discuss the theoretical expectations of misalignments in isolated galaxies. We find that the lenses requiring a weak external shear, even though some do have a known environment, are not misaligned, consistent with theoretical considerations.

| - Figure 2 is not terribly useful. None of the panels is independent, and it
|   appears from the error bars that most of the points that were computed using
|   a larger radius are not much different from the points that were computed
|   using a smaller radius. While it would be terrific to know how the ellipticity
|   and alignments vary with distance from the lens, that question needs to be
|   answered with independent data points (i.e., independent annuli in the
|   analysis) and I don’t think the authors’ data is sufficient for the task.

As the inertia tensor depends on the square of the radius, changes in the estimated shape are conceivable when changing the radius of enclosure (different multiples of R_e). Thus, this Figure serve to some degree to assess the robustness of the estimated shape.

While we agree with the referee that our data is not sufficient to conclusively identify the trend of the increasing ellipticity of the stellar distribution with increasing radius we observe, it is interesting to find out whether it persists.

| - The authors posit that 0712 and 2016 are post merger-systems, but that seems
|   to be based solely on their lensing reconstructions and nothing else. Such an
|   extrapolation from one known post-merger lensing system with similar properties
|   (1608) doesn’t seem warranted.

We agree with the referee that the paper in its previous form may have put too much emphasis on this statement. We make now clear in the Conclusions, where we reiterate on our results, that this is a speculation based on the close similarities to the known merger system 1608. We have removed reference to this from the abstract.